\documentclass{beamer}

\usepackage{amsmath,amsthm,amssymb,enumerate}

\begin{document}

% what is dna
\begin{frame}

\end{frame}

% using dna to encode strings

% biological operations on dna

% logical operations on strings

% the dna computing model (restricted? unrestricted?)
\begin{frame}
\frametitle{The Restricted Model}
A \emph{tube} is a multiset of strings over some alphabet $\Sigma$. We have the following operations on tubes:
\begin{itemize}
\item \emph{Separate}. Given a tube $T$ and a symbol $s \in \Sigma$, this produces two tubes $+(T,s)$ and $-(T,s)$ where $+(T,s)$ is the set of strings of $T$ which contain the symbol $s$, and $-(T,s)$ is the set of strings of $T$ which do not contain $s$. The original tube $T$ is destroyed.
\item \emph{Merge}. Given tubes $T_1,T_2$, this produces $T_1 \cup T_2$. $T_1, T_2$ are destroyed.
\item \emph{Detect}. Given a tube $T$, produces ``yes'' if $T$ is nonempty, and ``no'' otherwise.
\end{itemize}
\end{frame}

\begin{frame}

introduce program notation
\end{frame}

\begin{frame}
\frametitle{The Unrestricted Model}

The operations on the previous slide give what is called the \emph{restricted model} of DNA computing. The \emph{unrestricted model} is given by including the operation

\emph{Amplify}. Which, given a tube $T$, produces two new tubes $T_1, T_2$ such that $T_1 = T_2 = T$, and destroys $T$.
\end{frame}

% example: deciding whether or not a graph is 3-colourable (NP-complete problem)
\begin{frame}
\frametitle{Example: A DNA program for the 3-colourability problem}

Given an $n$ vertex graph $G$ with edges $e_1,\ldots,e_n$, let $\Sigma = \{r_1,b_1,g_1,r_2,b_2,g_2,\ldots,r_n,b_n,g_n\}$, and consider the following restricted program on input $T = \{\alpha | \alpha \subseteq \Sigma \wedge \alpha = c_1c_2,\ldots,c_n \wedge c_i \in \{r_i,b_i,c_i\}, i = 1,2,\ldots,n \}$

\begin{enumerate}[(1)]
\item Input$(T)$.
\item for each edge $e_k \in \{e_1,\ldots,e_z\}$, let $e_k = \langle i,j \rangle$ in
  \begin{enumerate}[(i)]
  \item $T_{red} \leftarrow +(T,r_i)$ and $T_{blue \vee green} \leftarrow -(T,r_i)$
  \item $T_{blue} \leftarrow +(T_{blue \vee green},b_i)$ and $T_{green} \leftarrow -(T_{blue \vee green},b_i)$
  \item $T_{red}^{good} \leftarrow -(T_{red},r_j)$
  \item $T_{blue}^{good} \leftarrow -(T_{blue},b_j)$
  \item $T_{green}^{good} \leftarrow -(T_{green},g_j)$
  \item $T' = T_{red}^{good} \cup T_{blue}^{good}$
  \item $T = T_{gree}^{good} \cup T'$
  \end{enumerate}
\item Output(Detect$(T)$).
\end{enumerate}
\end{frame}

% measures of computational complexity for dna computing
\begin{frame}
\frametitle{Measures of Complexity for DNA Programs}

The number of biological steps performed gives a straightforward measure of time complexity, and the highest number of strands of DNA present at any step during a DNA computaiton can be thought of as it's space complexity. Therefore, we consider the number of \emph{bio steps} a DNA program takes, and the number of \emph{strands} it requires, as a function of the input.

Note that in the restricted model, the number of strands required is number of strands in the input test tube, as no operation creates more strands.
\end{frame}

\begin{frame}
\frametitle{Example: Complexity of the program for 3-colourability}

For a graph with $n$ vertices and $k$ edges \ldots

The number of strands required is $\lvert T \rvert = 3^n$, one strand for each possible assignment of colours to vertices.

The number of bio steps required is $7k + 2 \in \mathcal{O}(k)$, as we perform steps $(i - vii)$ once for each edge, and a constant number of other steps. (Alternatively, $\mathcal{O}(n^2)$ as $k \leq n^2$).
\end{frame}

% table of results / complexity from Boneh "on the computational power of DNA"?

\begin{frame}

we seem to have solved an NP-complete problem (3-colourability) in time polynomial in the size of the graph!

(problems: the amount of DNA still grows exponentially, simulation of DNA computation on a TM does not preserve this due to the high degree of parallelism present in the separate operation)

\end{frame}

% the model is turing complete (winfree) 

\begin{frame}
It is easy to see that we can simulate both the restricted and unrestricted model with a Turing machine. (how to simulate other direction?) (winfree shows that only annealing and ligation are needed, must explain this earlier)

This is done by proving that we can simulate an arbitrary \emph{blocked cellular automaton} (BCA) with a DNA program. Since BCAs are capable of universal computation, so then are DNA programs.

\end{frame}

% do we really want to go in to this?

% complexity / BP / NBP (really?)

% error rates and the amplify operation

We have been assuming that our biological operations are error free. In reality, they are not. Further, DNA strands are prone to dissolving, and errors in the annealing process sometimes occur.

In the universal model, we can address this by using the amplify operation every so many steps to double the amount of DNA, and by double-encoding our strings in the starting test tube. (that is, instead of $s_1,\ldots,s_n$, encode $s_1,\ldots,s_n,s_1,\ldots,s_n$.

% real-world dna computing (adleman, the DES team, and impracticality)



% other biocomputing models 

Deoxyribozyme-based. (tictactoe)

Programmable automata that can interact with the surrounding microbiological environment.

\end{document}